% File french-msg.tex (messages issued by the FrenchPro package)
\msgheader{-\msgid-\space}
\msg{1}{the #1 macro already exists}{}
\msg{2}{#1 file not found.}{}
\msg{3}{\frenchpack\space is not active here\string!}{}
\msg{4}{entering now "Poor-Man-French-Style" way}{}
\msg{5}{the lettrine definition is wrong}{}
\msg{6}{check again the lettrine code}{}
\msg{7}{the lettrine is reduced to a lonely letter}{}
\msg{8}{\string\footnotetext{#1} lost %
	\MessageBreak %
        Code event. \string\protect\string\footnote}{}
\msg{9}{Corrupted/absent language.dat file}{}
\msg{10}{french package not loaded}{}
\msg{11}{applying the \string<\string< frpatch \string>\string> #1}{}
\msg{12}{frpatch.sty is out of date, delete the file}{}
\msg{13}{the character "#1" is already activated}{}
\msgheader{-13-\space}
\msg{13b}{double punctuation is then disactivated}{}
\msgheader{-\msgid-\space}
\msg{14}{closing not opened guillemets}{}
\msg{15}{the language \frenchname\space is numbered \the\language}{}
\msg{16}{the English language\space is numbered \the\language}{}
\msg{17}{\string\wrongtypedspaces\space not running %
              in this context}{}
\msg*{18}{\frenchname.sty is forcing the \string\nofrenchguillemets\space %
          option in math with AmSLaTeX.}{}
\msg{19}{using the internal language number \the\language}{}
\msg*{20}{WARNING\string: the French language is undefined in your format.}{}
\msg*{20b}{\frenchname.sty\string: the French language is undefined %
           (ERROR\string!)}{}
\msg{21}{#1 is not defined}{}
\msg{22}{abbreviation of \string"#1\string" not found}{}
\msg*{23}{Package\string: \frenchpack\space\frenchstyleid%
          \space(B.Gaulle)}{} 
\msg{24}{\frenchname.sty is using, in this document, the font encoding %
         \f@encoding.}{}%
\msg{25}{\frenchname.sty is issuing his messages in %
         \ifEightBitOutput8bits.\else7bits mode %
                    "a la TeX".\fi %
         ^^J}{} 
\msg{26}{Error found in \frenchname.sty\string! %
         \MessageBreak %
         (look at e.g. the language.dat file)}{}
\msg*{27}{\frenchname.sty\string: language \l@no\space (#1) was initially %
           (at initex) numbered \csname l@#1\endcsname\space(ERROR\string!)}{}
\msg{28}{WARNING\string: TeX Version 2 don't allow to use the %
         8bits chars with diacritics}{}
\msg*{29}{***Warning***\string: TeX engine in use along with CM fonts %
          \MessageBreak %
          (as in current TeX format) isn't sufficient to hyphenate %
          \MessageBreak %
           words containing diacritics (like in French).}{}
\msg{30}{ERROR\string! %
         Something wrong\string! your TeX engine can't hyphenate correctly %
         \MessageBreak %
         french words with diacritics *in that current font* or %
         encoding}{}
\msg{31}{Wrong French Hyphenation\string! %
         Are you sure to run with a format *********** %
         \MessageBreak %
         in which the french patterns were %
         installed at initex time\string? ***********}{}
\msg{32}{if yes try to switch to T1 font %
         encoding (\string\usepackage[T1]{fontenc})***}{}
\msg{33}{CHECK the list for hexadecimal TeX codes you can't display %
         \MessageBreak %
         and decide if they must be eliminated from \@kbencoding.kbc %
         \MessageBreak %
         assuming your ''locale'' settings are okay on your system.}{}
\msg{34}{this file and other auxiliary files require to %
          use the following %
         \MessageBreak %
         LaTeX packages\string: \frenchpack\string! %
         \MessageBreak %
         check \string\usepackage\space or remove these files. %
         Typesetting is aborted\string!}{}
\msg{35}{kbconfig\string: Redefining \string ^\string ^\string Y uccode. %
         \MessageBreak %
         kbconfig\string: Redefining \string ^\string ^\string Z uccode.}{}
\msg{36}{ERROR\string: invalid name of LaTeX format %
          (\string\fmtname). %
         \MessageBreak %
         CORRECT it to\string: \string\fmtname=LaTeX2e\space %
         or ancient name.}{}
\msg{36b}{ERROR\string: non standard LaTeX format %
                       (\string\fmtname)}{}
\msg{37a}{ERROR\string: the french package is already active here.}{}
\msg{37b}{ERROR\string: the french package is stll not active here.}{}
\msg{38}{The language #1 is used as language number \the\workl@.}{}
\msg{39}{Error\string: file language.dat not found, trying to load %
         US-english hyphenation file}{}
\msg*{40}{Writing kb8to7.lex and kb7to8.lex}{}
\msg{41}{your format is out of date, please run initex again\string!}{}
\msg{42}{The \frenchname\space patch file (frpatch.sty) is not suitable %
         \MessageBreak %
         for this version of the "\frenchpack" package dated \FSfd}{}
\msg{43}{french.all is LOADED^^J}{}
\msg{44}{ERROR the command \LaMacro\space is undefined\string!}{}
\msg{45}{executing \LaMacro}{}
\msg{46}{This commande is already defined in \frenchname.sty %
         \MessageBreak %
         Possible problem with \frenchpack\space (\Aboi #1 macro)}{}
\msg*{47}{i must stop because your TeX engine is unable to %
          generate 8bit output codes, sorry\string!}{}
\msg{48}{Reading the configuration file of \frenchpack}{}
\msg{49}{closing the guillemets too early}{}
\msg{50}{Error\string: the french language is undefined in %
                       language.dat file.}{}
\msg{51}{ERROR\string: this document was not converted to 8bit...}{}
\msg*{52}{Error\string: the \frenchpack\space package doesn't run in %
          such minimal document class, sorry\string!}{}
\msg{53}{the guillemets environment is not usable with %
         \MessageBreak %
         the \string\noeveryparguillemets\space option}{}
\msg{54}{ERROR\string: \string\unusedslot\space (\unusedslot) is invalid}{}
\msg{55}{ERROR \string\unusedslot\space character %
              (\@unusedslot) is not activated.}{}
\msg*{56}{Processed 8bits characters are\string:}{}
\msg{57}{keyboard package #1 loaded for the font encoding % 
              \f@encoding.^^J}{}
\msg{58}{value of #1 ignored}{}
\msg*{59}{FATAL ERROR\string: unusual strange text generated.}{}
\msg{60}{dot missing after \string\etc}{}
\msg{61}{hyph\hyphchar.tex file not found.}{}
\msg{62}{additional hyphenation file loading %
         for \hyphchar\space fonts with specific hyphen char}{}
\msg{63}{french.sty or frenchle.sty files not found. %
         \MessageBreak %
         Check if files exist somewhere in your system.}{}
\msg{64}{ERROR\string: invalid \string\fmtname\space in lplain.tex}{}
\msg{65}{\frenchname.sty is loading the translations %
         for bibliography \string:}{}
\msg*{66}{ERROR\string! This file can't be typeset without any input %
          \MessageBreak %
          encoding declaration (look at keyboard or inputenc packages.)}{}
\msg*{67}{WARNING it seeems your are using inputenc and keyboard, %
         please chose\string!}{}
\msg*{68}{ERROR\string: \frenchpack\space is no more running %
          with 2.09 emulation, sorry\string!}{}
\msg{69}{kbconfig\string: Intercepting the dump control sequence for %
                   eliminating few invalid chars.}{}
\msg{70}{kbconfig\string: modifying \string\@tabacckludge\space for math.}{}
\msg{71}{WARNING\string: if babel is used,%
         \MessageBreak
         specify option \frenchname(pro/le)}{}
\msg{72}{kbconfig\string: utf8 encoding can't be completed with plain\string! %
	 \MessageBreak %
         kbconfig\string: continuing with no Unicode characters.}{}
\msg*{73}{\frenchname.sty was loaded too early\string!}{}
\msg*{74}{ERROR, keyboard package loaded after %
                 \frenchpack!}{}
\msg{75}{kbconfig\string: utf8 encoding not found via inputenc %
                          package\string! %
	 \MessageBreak %
         kbconfig\string: continuing with no Unicode characters.}{}
\msg{76}{kbconfig\string: no language dependencies available within %
                          Plain TeX\string!}{}
\msg{77}{No #1\space characters defined for English.}{}
\msg*{78}{ERROR\string: noutf8 option clash.}{}
\msg*{79}{keyboard.sty\string: no local option given, %
                               no utf8 coding.}{}
\msg{80}{Babel is willing to load the \CurrentOption\space option %
         \MessageBreak but the mlp package is already loaded, %
         please choose %
         \MessageBreak either mlp with \CurrentOption\space option, %
         \CurrentOption\space alone %
         \MessageBreak or babel with \CurrentOption\space option}%
         {\help{just remove the superfluous package or option.}}
\msg*{81}{WARNING\string: "msg" package not found\string; %
          \MessageBreak continuing without message texts}{}
\msg{82}{command \LaMacro\space is not defined now}{}
\msg{83}{Testing macros before \string\begin{document}^^J}{}
\msg{84}{Now testing again but AFTER \string\begin{document}^^J}{}
\msg{85}{Warning\string: the fixlanguage option %
         was not provided at the babelbib call}{}
\msg{86}{keyboard.sty\string: %
         The mltex feature is nullified due to the #1\space option.}{}
\msg{87}{ERROR\string: % 
        \MessageBreak %
         babelbib should only be used with babel}{}
\msg{88}{No characters defined %
         \MessageBreak %
         by keyboard encoding change to `utf8'\MessageBreak}{}
\msg{89}{kbconfig\string: Keyboard character used is undefined\MessageBreak %
         in keyboard encoding `utf8'}%
       {You need to provide a  definition with %
        \noexpand\DeclareInputText\MessageBreak or %
        \noexpand\DeclareInputMath before using this key.}
\msg{90}{Invalid Unicode char value \string < 00A0}{\@eha}
\msg{91}{Unicode char #1 not set up\space %
         for use with LaTeX}{\@eha}
\msg{92}{Yous should either\string:\MessageBreak %
         - use an engine with the MlTeX option, or\MessageBreak %
         - switch to font encoding T1 as default.}{}
\msg*{93}{You are using two French options for Babel, %
          \MessageBreak %
          please choose only one.}{}
\msgheader{-**-\space}
\msg*{**a}{\frenchpack\space serial number is *\@FrVal*.}{}
\msg*{**b}{\@FrVal\space of \frenchpack.}{}
\msgheader{}
\msg{*}{the message \msgid\space is unknown in the \frenchname\space %
        package}{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

